---
title: "Metro DID: Minimum Wage Pass-Through into FAFH CPI"
output:
  html_document:
    toc: true
    toc_depth: 2
---

We build  a metro-month DID panel for minimum-wage pass-through into Food Away From Home CPI. We combine the uploaded Los Angeles FAFH workbook, uploaded state and local minimum-wage files, metro FAFH CPI series for New York and Chicago from FRED, and a national overall CPI control.

## 1. Packages

```{r setup, message=FALSE, warning=FALSE}
req_pkgs <- c(
  "readxl", "readr", "dplyr", "tidyr", "stringr", "lubridate",
  "purrr", "tibble", "ggplot2", "quantmod", "fixest", "zoo"
)

missing_pkgs <- req_pkgs[!vapply(req_pkgs, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_pkgs) > 0) install.packages(missing_pkgs)

invisible(lapply(req_pkgs, library, character.only = TRUE))
```

## 2. File paths

```{r paths}
la_fafh_path  <- "SeriesReport-20260406194017_6fa512.xlsx"
state_mw_path <- "state_minimum_wage_hikes_since_2009.csv"
local_mw_path <- "local_minimum_wage_hikes_since_2009.csv"
national_fafh_path <- "1978 to 2025 CPI FAFH Seasonally Adjusted.xlsx"
```

## 3. Helper functions

```{r helpers}
read_bls_xlsx_monthly <- function(path) {
  raw <- readxl::read_excel(path, sheet = 1, skip = 11)

  month_order <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                   "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

  raw %>%
    dplyr::select(Year, dplyr::all_of(month_order)) %>%
    dplyr::filter(!is.na(Year)) %>%
    tidyr::pivot_longer(
      cols = dplyr::all_of(month_order),
      names_to = "month_abb",
      values_to = "value"
    ) %>%
    dplyr::mutate(
      year = as.integer(Year),
      month = match(month_abb, month_order),
      date = as.Date(sprintf("%04d-%02d-01", year, month)),
      value = suppressWarnings(as.numeric(value))
    ) %>%
    dplyr::arrange(date) %>%
    dplyr::select(date, value)
}

get_fred_series <- function(series_id, value_name = "value") {
  xt <- quantmod::getSymbols(Symbols = series_id, src = "FRED", auto.assign = FALSE)
  tibble::tibble(
    date = as.Date(zoo::index(xt)),
    value = as.numeric(xt[, 1])
  ) %>%
    dplyr::mutate(date = as.Date(format(date, "%Y-%m-01"))) %>%
    dplyr::rename(!!value_name := value)
}

parse_rate_num <- function(x) {
  x <- as.character(x)
  nums <- stringr::str_extract_all(x, "[0-9]+\\.?[0-9]*")
  purrr::map_dbl(nums, function(v) {
    if (length(v) == 0) return(NA_real_)
    max(as.numeric(v), na.rm = TRUE)
  })
}

month_floor_after_cutoff <- function(date_vec, cutoff_day = 24L) {
  d <- as.Date(date_vec)
  shifted <- lubridate::floor_date(d, unit = "month")
  idx <- !is.na(d) & lubridate::day(d) >= cutoff_day
  shifted[idx] <- lubridate::ceiling_date(d[idx], unit = "month")
  as.Date(shifted)
}

build_monthly_rate_series <- function(changes_df, entity_col, eff_col, from_col, to_col,
                                      start_date, end_date) {
  entity_sym <- rlang::sym(entity_col)
  eff_sym    <- rlang::sym(eff_col)
  from_sym   <- rlang::sym(from_col)
  to_sym     <- rlang::sym(to_col)

  months <- tibble::tibble(date = seq(as.Date(start_date), as.Date(end_date), by = "month"))
  entities <- changes_df %>% dplyr::distinct(!!entity_sym) %>% dplyr::pull(!!entity_sym)

  purrr::map_dfr(entities, function(entity_value) {
    df <- changes_df %>%
      dplyr::filter(!!entity_sym == entity_value) %>%
      dplyr::arrange(!!eff_sym)

    baseline <- df %>% dplyr::slice(1) %>% dplyr::pull(!!from_sym)
    baseline <- ifelse(length(baseline) == 0 || is.na(baseline[1]), NA_real_, baseline[1])

    months %>%
      dplyr::left_join(
        df %>% dplyr::select(date = !!eff_sym, rate = !!to_sym),
        by = "date"
      ) %>%
      dplyr::arrange(date) %>%
      tidyr::fill(rate, .direction = "down") %>%
      dplyr::mutate(
        rate = ifelse(is.na(rate), baseline, rate),
        !!entity_col := entity_value,
        .before = date
      )
  })
}

lag_within  <- function(x, k = 1L) dplyr::lag(x, n = k)
lead_within <- function(x, k = 1L) dplyr::lead(x, n = k)
```

## 4. Metro outcome series

Read Los Angeles from CPI FAFH,
Pull New York and Chicago from FRED.

```{r outcomes}
la_outcome <- read_bls_xlsx_monthly(la_fafh_path) %>%
  dplyr::mutate(metro = "Los Angeles") %>%
  dplyr::rename(fafh = value)

ny_outcome <- get_fred_series("CUURA101SEFV", "fafh") %>%
  dplyr::mutate(metro = "New York")

chi_outcome <- get_fred_series("CUURA207SEFV", "fafh") %>%
  dplyr::mutate(metro = "Chicago")

outcome_panel <- dplyr::bind_rows(la_outcome, ny_outcome, chi_outcome) %>%
  dplyr::filter(date >= as.Date("2009-01-01"), !is.na(fafh))
```

## 5. Overall CPI control

Pull the common inflation control series

```{r cpi-control}
cpi_all_us <- get_fred_series("CPIAUCNS", "cpi_all_us") %>%
  dplyr::filter(date >= as.Date("2009-01-01"))
```

## 6. Minimum-wage input tables

Read the state and local wage-change files and convert them into clean monthly change tables

```{r mw-inputs}
state_mw_raw <- readr::read_csv(state_mw_path, show_col_types = FALSE)
local_mw_raw <- readr::read_csv(local_mw_path, show_col_types = FALSE)

state_changes <- state_mw_raw %>%
  dplyr::transmute(
    state = jurisdiction,
    effective_month = as.Date(effective_date_assumption),
    from_rate = as.numeric(from_rate),
    to_rate   = as.numeric(to_rate)
  ) %>%
  dplyr::filter(!is.na(effective_month), !is.na(from_rate), !is.na(to_rate)) %>%
  dplyr::arrange(state, effective_month)

local_changes <- local_mw_raw %>%
  dplyr::transmute(
    jurisdiction = jurisdiction,
    effective_date = as.Date(effective_date),
    effective_month = month_floor_after_cutoff(effective_date, cutoff_day = 24L),
    from_rate = parse_rate_num(before_rate_raw),
    to_rate   = parse_rate_num(to_rate_raw)
  ) %>%
  dplyr::filter(!is.na(effective_month), !is.na(to_rate)) %>%
  dplyr::group_by(jurisdiction, effective_month) %>%
  dplyr::summarise(
    from_rate = max(from_rate, na.rm = TRUE),
    to_rate   = max(to_rate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    from_rate = ifelse(is.infinite(from_rate), NA_real_, from_rate),
    to_rate   = ifelse(is.infinite(to_rate), NA_real_, to_rate)
  ) %>%
  dplyr::arrange(jurisdiction, effective_month)

federal_changes <- tibble::tribble(
  ~entity,     ~effective_month, ~from_rate, ~to_rate,
  "Federal",  as.Date("2009-01-01"), 6.55, 6.55,
  "Federal",  as.Date("2009-08-01"), 6.55, 7.25
)

sample_start <- as.Date("2009-01-01")
sample_end   <- max(outcome_panel$date, na.rm = TRUE)

federal_monthly <- build_monthly_rate_series(
  changes_df = federal_changes,
  entity_col = "entity",
  eff_col    = "effective_month",
  from_col   = "from_rate",
  to_col     = "to_rate",
  start_date = sample_start,
  end_date   = sample_end
) %>%
  dplyr::filter(entity == "Federal") %>%
  dplyr::select(date, federal_rate = rate)

state_monthly <- build_monthly_rate_series(
  changes_df = state_changes,
  entity_col = "state",
  eff_col    = "effective_month",
  from_col   = "from_rate",
  to_col     = "to_rate",
  start_date = sample_start,
  end_date   = sample_end
)

local_monthly <- build_monthly_rate_series(
  changes_df = local_changes,
  entity_col = "jurisdiction",
  eff_col    = "effective_month",
  from_col   = "from_rate",
  to_col     = "to_rate",
  start_date = sample_start,
  end_date   = sample_end
)
```

## 7. Metro mapping

Map each metro to the state and local jurisdictions used to build treatment

```{r metro-mapping}
metro_core_state <- tibble::tribble(
  ~metro,         ~core_state,
  "Los Angeles", "California",
  "New York",    "New York",
  "Chicago",     "Illinois"
)

metro_state_components <- tibble::tribble(
  ~metro,         ~state,
  "Los Angeles", "California",
  "New York",    "New York",
  "New York",    "New Jersey",
  "New York",    "Pennsylvania",
  "Chicago",     "Illinois",
  "Chicago",     "Indiana",
  "Chicago",     "Wisconsin"
)

metro_local_components <- tibble::tribble(
  ~metro,         ~jurisdiction,
  "Los Angeles", "Los Angeles County",
  "Los Angeles", "Los Angeles, CA",
  "Los Angeles", "Santa Monica, CA",
  "Los Angeles", "Malibu, CA",
  "Chicago",     "Cook County, IL"
)
```

## 8. Metro-month treatment measures

Build the baseline binding wage and the multistate sensitivity measure

```{r treatment-panel}
metro_core_state_monthly <- metro_core_state %>%
  dplyr::left_join(state_monthly, by = c("core_state" = "state")) %>%
  dplyr::rename(core_state_rate = rate)

metro_local_monthly <- metro_local_components %>%
  dplyr::left_join(local_monthly, by = "jurisdiction") %>%
  dplyr::group_by(metro, date) %>%
  dplyr::summarise(local_rate_max = max(rate, na.rm = TRUE), .groups = "drop") %>%
  dplyr::mutate(local_rate_max = ifelse(is.infinite(local_rate_max), NA_real_, local_rate_max))

metro_multistate_monthly <- metro_state_components %>%
  dplyr::left_join(state_monthly, by = "state") %>%
  dplyr::group_by(metro, date) %>%
  dplyr::summarise(multistate_avg_rate = mean(rate, na.rm = TRUE), .groups = "drop")

metro_mw_panel <- outcome_panel %>%
  dplyr::distinct(metro, date) %>%
  dplyr::left_join(federal_monthly, by = "date") %>%
  dplyr::left_join(
    metro_core_state_monthly %>% dplyr::select(metro, date, core_state_rate),
    by = c("metro", "date")
  ) %>%
  dplyr::left_join(metro_local_monthly, by = c("metro", "date")) %>%
  dplyr::left_join(metro_multistate_monthly, by = c("metro", "date")) %>%
  dplyr::mutate(
    core_state_rate     = pmax(core_state_rate, federal_rate, na.rm = TRUE),
    mw_core_local       = pmax(core_state_rate, local_rate_max, na.rm = TRUE),
    mw_core_local       = ifelse(is.infinite(mw_core_local), core_state_rate, mw_core_local),
    multistate_avg_rate = pmax(multistate_avg_rate, federal_rate, na.rm = TRUE)
  )
```

## 9. Regression panel

Join outcomes, treatment, and controls, then construct log changes and treatment indicators

```{r panel-build}
panel <- outcome_panel %>%
  dplyr::left_join(cpi_all_us, by = "date") %>%
  dplyr::left_join(metro_mw_panel, by = c("metro", "date")) %>%
  dplyr::arrange(metro, date) %>%
  dplyr::group_by(metro) %>%
  dplyr::mutate(
    log_fafh            = log(fafh),
    log_cpi_all_us      = log(cpi_all_us),
    log_mw_core_local   = log(mw_core_local),
    log_mw_multistate   = log(multistate_avg_rate),
    dlog_fafh           = log_fafh - lag_within(log_fafh),
    dlog_cpi_all_us     = log_cpi_all_us - lag_within(log_cpi_all_us),
    dlog_mw_core_local  = log_mw_core_local - lag_within(log_mw_core_local),
    dlog_mw_multistate  = log_mw_multistate - lag_within(log_mw_multistate),
    lead_dlog_mw_core   = lead_within(dlog_mw_core_local),
    lag_dlog_mw_core    = lag_within(dlog_mw_core_local),
    treated_core        = as.integer(!is.na(dlog_mw_core_local) & dlog_mw_core_local > 0),
    month               = factor(lubridate::month(date)),
    year                = factor(lubridate::year(date))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::filter(date >= as.Date("2009-02-01"))

readr::write_csv(panel, "metro_panel_cleaned.csv")
readr::write_csv(
  panel %>%
    dplyr::filter(treated_core == 1) %>%
    dplyr::select(metro, date, mw_core_local, core_state_rate, local_rate_max),
  "treated_months_by_metro.csv"
)

est <- panel %>%
  dplyr::filter(
    !is.na(dlog_fafh),
    !is.na(dlog_cpi_all_us),
    !is.na(dlog_mw_core_local),
    !is.na(dlog_mw_multistate)
  )
```

## 10. Main models

Estimate the baseline DID, a treated-month version, a dynamic version, and two robustness checks

```{r models}
m1 <- fixest::feols(
  dlog_fafh ~ dlog_mw_core_local + dlog_cpi_all_us | metro + month + year,
  data = est,
  cluster = ~ metro
)

m2 <- fixest::feols(
  dlog_fafh ~ treated_core + dlog_cpi_all_us | metro + month + year,
  data = est,
  cluster = ~ metro
)

m3 <- fixest::feols(
  dlog_fafh ~ lead_dlog_mw_core + dlog_mw_core_local + lag_dlog_mw_core + dlog_cpi_all_us |
    metro + month + year,
  data = est %>% dplyr::filter(!is.na(lead_dlog_mw_core), !is.na(lag_dlog_mw_core)),
  cluster = ~ metro
)

m4 <- fixest::feols(
  dlog_fafh ~ dlog_mw_multistate + dlog_cpi_all_us | metro + month + year,
  data = est,
  cluster = ~ metro
)

est <- est %>% dplyr::mutate(year_month = interaction(format(date, "%Y-%m"), drop = TRUE))

m5 <- fixest::feols(
  dlog_fafh ~ dlog_mw_core_local | metro + year_month,
  data = est,
  cluster = ~ metro
)
```

## 11. Results snapshot

Print the model summaries and implied elasticities

```{r results}
summary(m1)
summary(m2)
summary(m3)
summary(m4)
summary(m5)

beta1 <- coef(m1)["dlog_mw_core_local"]
beta4 <- coef(m4)["dlog_mw_multistate"]

cat(sprintf("Model 1 elasticity: %.4f -> 10%% MW increase implies %.3f%% FAFH increase\n", beta1, 10 * beta1))
cat(sprintf("Model 4 elasticity: %.4f -> 10%% MW increase implies %.3f%% FAFH increase\n", beta4, 10 * beta4))

if ("dlog_mw_core_local" %in% names(coef(m3))) {
  bsum <- sum(coef(m3)[c("lead_dlog_mw_core", "dlog_mw_core_local", "lag_dlog_mw_core")], na.rm = TRUE)
  cat(sprintf("Model 3 cumulative lead/current/lag elasticity: %.4f -> 10%% MW increase implies %.3f%% FAFH increase\n", bsum, 10 * bsum))
}
```

## 12. Export tables and plot

Write outputs into txt and create event study plot png (var p)

```{r exports}
broom_fixest <- function(model, model_name) {
  ct <- as.data.frame(fixest::coeftable(model))
  ct$term <- rownames(ct)
  rownames(ct) <- NULL
  names(ct)[1:4] <- c("estimate", "std_error", "t_value", "p_value")
  tibble::as_tibble(ct) %>%
    dplyr::mutate(model = model_name) %>%
    dplyr::select(model, term, estimate, std_error, t_value, p_value)
}

coef_table <- dplyr::bind_rows(
  broom_fixest(m1, "m1_core_local"),
  broom_fixest(m2, "m2_treated_dummy"),
  broom_fixest(m3, "m3_dynamic"),
  broom_fixest(m4, "m4_multistate_avg"),
  broom_fixest(m5, "m5_yearmonth_fe")
)
readr::write_csv(coef_table, "model_coefficients.csv")

model_text <- capture.output({
  cat("\n==============================\n")
  cat("MODEL 1: Main intensity DID (core state + mapped local ordinances)\n")
  print(summary(m1))

  cat("\n==============================\n")
  cat("MODEL 2: Binary treated-month DID\n")
  print(summary(m2))

  cat("\n==============================\n")
  cat("MODEL 3: Dynamic lead/current/lag DID\n")
  print(summary(m3))

  cat("\n==============================\n")
  cat("MODEL 4: Sensitivity using equal-weight multistate treatment\n")
  print(summary(m4))

  cat("\n==============================\n")
  cat("MODEL 5: Robustness with metro FE + year-month FE\n")
  print(summary(m5))

  cat("\n==============================\n")
  cat("IMPLIED PASS-THROUGH FOR A 10% MINIMUM-WAGE INCREASE\n")
  cat(sprintf("Model 1 elasticity: %.4f -> 10%% MW increase implies %.3f%% FAFH increase\n", beta1, 10 * beta1))
  cat(sprintf("Model 4 elasticity: %.4f -> 10%% MW increase implies %.3f%% FAFH increase\n", beta4, 10 * beta4))

  if ("dlog_mw_core_local" %in% names(coef(m3))) {
    bsum <- sum(coef(m3)[c("lead_dlog_mw_core", "dlog_mw_core_local", "lag_dlog_mw_core")], na.rm = TRUE)
    cat(sprintf("Model 3 cumulative lead/current/lag elasticity: %.4f -> 10%% MW increase implies %.3f%% FAFH increase\n", bsum, 10 * bsum))
  }
})

writeLines(model_text, con = "did_model_summary.txt")

plot_df <- est %>%
  dplyr::select(date, metro, dlog_fafh, dlog_mw_core_local, treated_core) %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(
    mean_dlog_fafh = mean(dlog_fafh, na.rm = TRUE),
    treated_share = mean(treated_core, na.rm = TRUE),
    .groups = "drop"
  )

p <- ggplot2::ggplot(plot_df, ggplot2::aes(x = date, y = mean_dlog_fafh)) +
  ggplot2::geom_line() +
  ggplot2::geom_point(data = subset(plot_df, treated_share > 0), size = 2) +
  ggplot2::labs(
    title = "Average metro FAFH inflation with treated months marked",
    subtitle = "Dots indicate months where at least one metro experiences a minimum-wage increase",
    x = NULL,
    y = "Average monthly log change in FAFH CPI"
  ) +
  ggplot2::theme_minimal()

ggsave("event_study_plot.png", plot = p, width = 10, height = 5, dpi = 300)
```

## Final note

The notebook writes the cleaned panel, treated-month file, model summaries, coefficient table, and the plot into the working directory.
